---
title: "Study on COVID Death Rate in Nursing Home"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mice)
library(rpart)
library(rpart.plot)
library(Amelia)
library(missForest)
library(doParallel)
library(nnet)
library(MASS)
```

SD: Submitted Data
PQA: Passed Quality Assurance Check
RTA: Residents Total Admissions COVID-19
NB: Number of All Beds
OB: Total Number of Occupied Beds
CTR: Able to Test or Obtain Resources to Test All Current Residents Within Next 7 Days
CTS: Able to Test or Obtain Resources to Test All Staff and/or Personnel Within Next 7 Days
TR: COVID-19 Point-of-Care Tests Performed on Residents Since Last Report
TS: COVID-19 Point-of-Care Tests Performed on Staff and/or Personnel Since Last Report
STC: Staff Total Confirmed COVID-19
STD: Staff Total COVID-19 Deaths
SN: Shortage of Nursing Staff
SC: Shortage of Clinical Staff
SA: Shortage of Aides
SO: Shortage of Other Staff
RCR: Total Resident Confirmed COVID-19 Cases Per 1,000 Residents
RDR: Total Resident COVID-19 Deaths Per 1,000 Residents
NRNI: Number of Residents with New Influenza
NRAR: Number of Residents with Acute Respiratory Illness Symptoms Excluding COVID-19 and/or Influenza
ATT: During Past Two Weeks Average Time to Receive COVID-19 Test Results from Non-Point-of-Care Tests
NSRA: Number of Staff and/or Personnel with Acute Respiratory Illness Symptoms Excluding COVID-19 and/or Influenza
ABHR: Alcohol-Based Hand Rub (ABHR) Available
N95RS: N95 Respirator Strategy for Optimization
FMS: Face Masks Strategy for Optimization
EPS: Eye Protection Strategy for Optimization
GS: Gowns Strategy for Optimization
GLS: Gloves Strategy for Optimization


# Data Exploration and cleanup

```{R}
# Import dataset
data <- read.csv("COVID_19_Nursing_Home_Data.csv", head=TRUE)
```


```{R}
# Cleanup
data[data == "" | data == " "] <- NA
# Drop subjects that did not submit data in this cycle or data did not pass
# QA check
# Drop rows where response is missing
data = data[(data$SD == "Y" | data$PQA == "Y") & !is.na(data$RDR), ]
# Drop submit data and QA check status
data <- subset(data, select = -c(SD, PQA))
# Remove subjects with only NA values
data = data[!!rowSums(!is.na(data)),]
# Use bed occupation rate instead of bed counts
data$BOC = data$OB / data$NB
# Drop bed counts
data <- subset(data, select = -c(OB, NB))
name <- names(data)

for (i in 1:length(name)) {
  col =name[i]
  if (class(data[, col]) == "character") {
    data[, col]= as.factor(data[, col])
  }
}

# Check which columns are missing
index <- names(data)
tabcol2 <- rep(NA, length(index))
for (i in 1:length(index)){
  col = index[i]
  tabcol2[i] = length(which(is.na(data[,col])))
}
index[which(tabcol2/length(data[,1])>0.05)] # missing percentage > 0.05
index[which(tabcol2/length(data[,1])>0)] # missing data

# Use decision tree to find significant variate that we want to keep
df <- rpart(data$RDR~ .,
            data = data,
            control = rpart.control(minsplit = 1,
                                    minbucket = 1,
                                    maxdepth = 5,
                                    cp = 0,
                                    xval = 6))
rpart.plot(df)

# variables to consider
var_import <- names(df$variable.importance)
```

```{r fig.width = 5, fig.height = 5}
# Visualize missing pattern of original dataset
md.pattern(data)
```

# Missing Mechanism:
```{r}
# Variable with missing values
var_miss <- index[which(tabcol2/length(data[,1])>0)]
# Only keep the significant ones we wanted to keep in previous section
var_miss <- intersect(var_miss, var_import)
# Significant variables that do not contain missing values
var_complete <- setdiff(var_import, var_miss)

# Compute missing mechanism of all variables with missing values
for (i in 1:length(var_miss)) {
  loopdata = data[var_import]
  misscol = var_miss[i]
  print(misscol)
  loopdata$R <- ifelse(is.na(data[,misscol]), 1, 0)
  loopdata[,misscol] <- NULL
  mechanism = glm(R ~ RCR + STC + RTA + STD,family = "binomial",
                  data=loopdata)
  print(summary(mechanism))
}
```

```{r}
# Visualization of missing pattern of dataset with only significant variables
data_import <- data[c(var_import, "RDR")]
md.pattern(data[var_import])
```


# MCAR Imputation
```{r}
# Variable MCAR, impute them with linear model
fit=glm(CTR~ RCR + STC + RTA + STD,data=data_import, family = "binomial")
data_import$impute.CTR=predict(fit,newdata=data_import)
data_import$impute.CTR[!is.na(data_import$CTR)]=
  data_import$CTR[!is.na(data_import$CTR)]
data_import$CTR = data_import$impute.CTR
data_import = subset(data_import, select = -c(impute.CTR))


fit=glm(CTS~ RCR + STC + RTA + STD,data=data_import, family = "binomial")
data_import$impute.CTS=predict(fit,newdata=data_import)
data_import$impute.CTS[!is.na(data_import$CTS)]=
  data_import$CTS[!is.na(data_import$CTS)]
data_import$CTS = data_import$impute.CTS
data_import = subset(data_import, select = -c(impute.CTS))

fit=glm(SN~ RCR + STC + RTA + STD,data=data_import, family = "binomial")
data_import$impute.SN=predict(fit,newdata=data_import)
data_import$impute.SN[!is.na(data_import$SN)]=
  data_import$SN[!is.na(data_import$SN)]
data_import$SN = data_import$impute.SN
data_import = subset(data_import, select = -c(impute.SN))

fit=glm(SA~ RCR + STC + RTA + STD,data=data_import, family = "binomial")
data_import$impute.SA=predict(fit,newdata=data_import)
data_import$impute.SA[!is.na(data_import$SA)]=
  data_import$SA[!is.na(data_import$SA)]
data_import$SA = data_import$impute.SA
data_import = subset(data_import, select = -c(impute.SA))

fit=glm(N95RS~ RCR + STC + RTA + STD,data=data_import, family = "binomial")
data_import$impute.N95RS=predict(fit,newdata=data_import)
data_import$impute.N95RS[!is.na(data_import$N95RS)]=
  data_import$N95RS[!is.na(data_import$N95RS)]
data_import$N95RS = data_import$impute.N95RS
data_import = subset(data_import, select = -c(impute.N95RS))

fit=multinom(EPS~ RCR + STC + RTA + STD, data = data_import)
data_import$impute.EPS=predict(fit,newdata=data_import)
data_import$impute.EPS[!is.na(data_import$EPS)]=
  data_import$EPS[!is.na(data_import$EPS)]
data_import$EPS = data_import$impute.EPS
data_import = subset(data_import, select = -c(impute.EPS))

fit=multinom(GS~ RCR + STC + RTA + STD, data = data_import)
data_import$impute.GS=predict(fit,newdata=data_import)
data_import$impute.GS[!is.na(data_import$GS)]=
  data_import$GS[!is.na(data_import$GS)]
data_import$GS = data_import$impute.GS
data_import = subset(data_import, select = -c(impute.GS))
```


```{r}
for (i in 1:length(names(data_import))) {
  col =names(data_import)[i]
  if (class(data_import[, col]) == "factor") {
    data_import[, col]= as.numeric(data_import[, col])
  }
}
```


# Non-MCAR Imputation

## EM algorithm
```{r}
set.seed(438)
data.imputed.em=amelia(data_import, m=5)
data.imputed.em <- data.imputed.em$imputations$imp5
```



# Analysis, Correlation


```{r}
### stage1
resp.zero <- which(data.imputed.em$RDR==0)
resp.nonzero <- which(data.imputed.em$RDR!=0)

data.imputed.em$bin.resp <- ifelse(data.imputed.em$RDR==0, 0, 1)

m.stage1 <- glm(bin.resp~RCR+STC+RTA+BOC+SN+GS+SA+TS+STD+EPS+N95RS+CTS+CTR,
                family=binomial(link="logit"),
                data=data.imputed.em)
summary(m.stage1)

### stage2
#### Box-cox transformation
bc <- boxcox(lm(RDR~RCR+STC+RTA+BOC+SN+GS+SA+TS+STD+EPS+N95RS+CTS+CTR,
                data=data.imputed.em[resp.nonzero,]))
lambda <- bc$x[which.max(bc$y)]

m.stage2 <- glm((RDR^lambda-1)/lambda~RCR+STC+RTA+BOC+SN+GS+SA+TS+STD+EPS+N95RS+CTS+CTR,
                data=data.imputed.em[resp.nonzero,])
summary(m.stage2)
```


```{r}
# stage 1 
library(knitr)
table1 <- data.frame(rbind(
  c("RCR",4.173e-03,  1.178e-04),
  c("STC",3.200e-02,  2.003e-03),
  c("RTA",3.267e-02,  2.587e-03),
  c("BOC",3.004e-01,  1.502e-01),
  c("TS",8.594e-04,  4.152e-04),
  c("STD",2.170e-01,  8.093e-02)))
colnames(table1) <- c("Predictor", "Estimates", "Standard Error")
kable(table1)

# stage 2 
library(knitr)
table2 <- data.frame(rbind(
  c("RCR",0.0026112,  0.0000462),
  c("STC",0.0052874,  0.0007866),
  c("RTA",0.0078496,  0.0003809),
  c("BOC",-1.8101680,  0.1144047),
  c("SA",0.2340079,  0.0880236),
  c("TS",-0.0014927,  0.0002382),
  c("STD",0.2139176,  0.0317822)))
colnames(table2) <- c("Predictor", "Estimates", "Standard Error")
kable(table2)

```


